Learning to Synthesize Rules for Rule-based Intrusion Detectors

### 2020-03-05 paper/paper.tex
pass on technique

\vspace{1ex}
\noindent\textbf{Overview.}~\tname\ takes on input one or more
examples of the same attack and produces on output a ranked list of
\emph{plausible rules}---\ie{}, rules that isolate the malicious
traffic produced by the attacks from the benign
traffic. Figure~\ref{fig:overview} shows the workflow of \tname{} as a
pipeline of three components, appearing numbered in the figure.
Public data sets of benign
traffic~\cite{tcpreplay,stratosphere-normal} and rules
exist~\cite{emerging-threats-open}. Consequently, a user of \tname{}
only needs to provide examples of attacks on input.

@@

\begin{figure}[t!]
  \centering
  \includegraphics[trim=0 395 50 0,clip,width=0.48\textwidth]{figs/nids-workflow}
  \vspace{-4ex}  
  \caption{The \tname\ workflow.\Mar{revisar}}
  \label{fig:overview}
  \vspace{-3ex}
\end{figure}

\tname\ works as follows. First, it creates a rule that captures the
malicious traffic produced by the examples provided on input. In this
step, \tname\ detects options directly from the input messages (see
Section~\ref{sec:rules-and-packets}). More precisely, it creates a
rule containing options that are guaranteed to be satisfied by the
malicious traffic. Let us refer to that rule as the \emph{seed rule}
and its corresponding set of options as \emph{seed options}. Note
that, as only the malicious traffic is considered in this step, the
seed rule is likely overspecified, \ie{}, it may be more restrictive
than the golden rule. Consequently, the use of that rule could result
in false negatives. Second, \tname\ generates alternative rules that
1) weaken the constraints expressed in the seed rule while 2) still
capturing the malicious traffic. Given that every option declared in a
rule needs to be satisfied for a rule to capture an attack, any subset
of the set of seed options, by definition, captures the attack
characterized by the malicious traffic. \tname{} leverages that
observation to search for plausible rules containing a proper subset
of options from the seed options. However, note that that approach
alone could result in false positives as it imposes no limit on to
what extent a the constraints induced from the rule can be
weakened. \tname{} uses benign traffic for that purpose. More
precisely, it searches for plausible rules derived from the seed rule
obtained in the previous step. Finally, \tname{} uses heuristic
functions obtained from a data set of existing rules to rank the
generated plausible rules by their similarity to existing
rules. Similarity is used to approximate user's intent. The following
sections detail each component of \tname.

>>>>

\vspace{1ex}
\noindent\textbf{Overview.}~\tname\ takes on input one or more
examples of an attack and produces on output a ranked list of
\emph{plausible rules}---\ie{}, rules that isolate the malicious
traffic produced by the attacks from the benign
traffic. Figure~\ref{fig:overview} shows the workflow of \tname{} as a
pipeline of three components, appearing numbered in the figure.
Public data sets of benign
traffic~\cite{tcpreplay,stratosphere-normal} and rules
exist~\cite{emerging-threats-open}. Consequently, a user of \tname{}
only needs to provide examples of attacks on input.

@@

\begin{figure}[t!]
  \centering
  \includegraphics[trim=0 340 50 0,clip,width=0.48\textwidth]{figs/nids-workflow}
  \vspace{-4ex}  
  \caption{The \tname\ workflow.}
  \label{fig:overview}
  \vspace{-3ex}
\end{figure}

\tname\ works as follows. First, it creates a rule that captures the
malicious traffic produced by the examples provided on input. In this
step, \tname\ detects options directly from the input messages (see
Section~\ref{sec:rules-and-packets}). More precisely, it creates a
rule containing options that are guaranteed to be satisfied by the
malicious traffic. Let us refer to that rule as the \emph{seed rule}
and its corresponding set of options as \emph{seed options}. Note
that, as only the malicious traffic is considered in this step, the
seed rule is likely overspecified, \ie{}, it may be more restrictive
than necessary. Consequently, the use of that rule could result in
false negatives. Second, \tname\ generates alternative rules that 1)
weaken the constraints expressed in the seed rule while 2) still
capturing the malicious traffic. Given that every option declared in a
rule needs to be satisfied for a rule to capture an attack, any subset
of the set of seed options, by definition, captures the attack
characterized by the malicious traffic. \tname{} leverages that
observation to search for plausible rules containing a proper subset
of options from the seed options. However, note that that approach
alone could result in false positives as it imposes no limit on the
extent the constraints induced from the rule can be weakened. \tname{}
uses benign traffic for that purpose. More precisely, it searches for
plausible rules derived from the seed rule obtained in the previous
step. Finally, \tname{} uses heuristic functions, obtained from a data
set of existing rules, to rank the generated plausible rules by their
similarity to existing rules. Similarity is used to approximate user's
intent. The following sections detail each component of \tname.
